PANCHANGE_data_preprocessing

### 2020-08-19 PANCHANGE_analysis.Rmd

@@

@@

```{r Disorder by Ethnicity}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Ethnicity) %>%
  count(Disorder)
```

@@

Questionnaire columns
```{r Questionnaire columns}

@@

```{r}
dat.long.gad.diff %>%
  ggplot(mapping = aes(y = gad.diff_score)) +
  geom_bar() +
  theme_minimal() +
  facet_wrap(facets = vars(Time_point))
```

```{r gather to long format Time_points}
dat.long.phq.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.diff_score",
    all_of(phq.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Time_point_raw,
    phq.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r}
dat.long.phq.diff %>%
  ggplot(mapping = aes(y = phq.diff_score)) +
  geom_bar() +
  theme_minimal() +
  facet_wrap(facets = vars(Time_point))
```

>>>>

Filter data frame to exclude participants that never answered COPING
```{r}

@@

Data availability OCI-R
```{r}

Descriptives of age
```{r}

@@

Ethnicity should be recoded
European
Mixed or multiplce ethnic origins
Africian or African British
Asian or Asian British
Arab
```{r Disorder by Ethnicity}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Ethnicity) %>%
  count(Disorder)
```

@@

@@

makes this into percentages
```{r}
dat.long.gad.diff %>%
  ggplot(mapping = aes(y = gad.diff_score)) +
  geom_bar() +
  theme_minimal() +
  facet_wrap(facets = vars(Time_point))
```

```{r gather to long format Time_points}
dat.long.phq.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.diff_score",
    all_of(phq.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Time_point_raw,
    phq.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.diff_score_", "")) %>%
  select(-Time_point_raw) %>%
  drop_na(phq.diff_score)

@@

```{r}
dat.long.phq.diff %>%
  filter(Time_point != "prepan_retro") %>%
  ggplot(mapping = aes(
    x = phq.diff_score,
    y = stat(density*width),
    fill = Time_point)) +
   geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
    ) +
  theme_minimal()
```

```{r}
dat.long.phq.diff %>%
  filter(Time_point == "prepan_retro") %>%
  ggplot(mapping = aes(
    x = phq.diff_score,
    y = stat(density*width),
    fill = Time_point)) +
   geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
    ) +
  theme_minimal()
```

### 2020-08-19 PANCHANGE_data_preprocessing2020-07-15.Rmd

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = TRUE
```

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

## Highest education (baseline, RAMP)
```{r highest level of education for ramp}
if(RAMP == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.what_is_your_highest_level_of_education == "PhD" ~ 4,
        employment.what_is_your_highest_level_of_education == "Postgraduate degree or equivalent" ~ 4,
        employment.what_is_your_highest_level_of_education == "Master's degree or equivalent" ~ 4,
        employment.what_is_your_highest_level_of_education == "Bachelor's degree or equivalent" ~ 4,
        employment.what_is_your_highest_level_of_education == "A-level or equivalent" ~ 3,
        employment.what_is_your_highest_level_of_education == "GCSE or equivalent" ~ 2,
        employment.what_is_your_highest_level_of_education == "None of these" ~ 1
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "None",
        `2` = "GCSE/CSE",
        `3` = "A-levels",
        `4` = "University")
    )

@@

Check highest education coding for RAMP
```{r Highest education check for RAMP}
if(RAMP == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.what_is_your_highest_level_of_education,
         employment.what_is_your_highest_level_of_education_numeric
  )
}
```

@@

```{r edgi employment and coping ocir}
if(EDGI == TRUE){
edgi.employment.coping.ocir <-  full_join(
    x = employment.raw.id,
    y = coping.ocir.raw.id,
    by = 
      c("Sample",
        "COPING_ID")
)
}
```

@@

Removing all prepandemic dataframes from NBR/RAMP (prepandemic measures don't exist for these cohorts)
```{r nbr/ramp data frame list}
if(NBR == TRUE | RAMP == TRUE) {
dfs.list <- list(
  dem.raw.id,
  mhd.raw.categories,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
}
```

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = TRUE
RAMP = FALSE
```

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(highest_education,
         highest_education_numeric,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

## Highest education (baseline, RAMP)
```{r highest level of education for ramp}
if(RAMP == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.what_is_your_highest_level_of_education == "PhD" ~ 4,
        employment.what_is_your_highest_level_of_education == "Postgraduate degree or equivalent" ~ 4,
        employment.what_is_your_highest_level_of_education == "Master's degree or equivalent" ~ 4,
        employment.what_is_your_highest_level_of_education == "Bachelor's degree or equivalent" ~ 4,
        employment.what_is_your_highest_level_of_education == "A-level or equivalent" ~ 3,
        employment.what_is_your_highest_level_of_education == "GCSE or equivalent" ~ 1,
        employment.what_is_your_highest_level_of_education == "None of these" ~ 5
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = NA_character_,
        `3` = "A-levels",
        `4` = "University",
        `5` = NA_character_)
    )

@@

Check highest education coding for RAMP
```{r Highest education check for RAMP}
if(RAMP == TRUE) {
employment.raw.id %>%
  select(highest_education,
         highest_education_numeric,
         employment.what_is_your_highest_level_of_education,
         employment.what_is_your_highest_level_of_education_numeric
  )
}
```

@@

```{r edgi employment and coping ocir}
if(EDGI == TRUE){
  dim(employment.raw.id)
  dim(coping.ocir.raw.id)

edgi.employment.coping.ocir <-  full_join(
    x = employment.raw.id,
    y = coping.ocir.raw.id,
    by = 
      c("Sample",
        "COPING_ID")
)

    dim(edgi.employment.coping.ocir)
}
```

@@

Removing all prepandemic dataframes from NBR/RAMP (prepandemic measures don't exist for these cohorts)
```{r nbr/ramp data frame list}
if(NBR == TRUE | RAMP == TRUE) {
dfs.list <- list(
  dem.raw.id,
  employment.raw.id,
  mhd.raw.categories,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
}
```
